---
title: "Reproducible Research Peer Assessment 1"
author: "Ning Guo"
date: "2016-8-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

download the date and read the data
```{r read data}
# set the file url 
fileurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
# create a temporary directory
td = tempdir()
# create the placeholder file
tf = tempfile(tmpdir=td, fileext=".zip")
# download into the placeholder file (curl method needed for Mac OS X)
download.file(fileurl, tf, method="curl")
# get the name of the first file in the zip archive
fname = unzip(tf, list=TRUE)$Name[1]
# unzip the file to the temporary directory
unzip(tf, files=fname, exdir=td, overwrite=TRUE)
# fpath is the full path to the extracted file
fpath = file.path(td, fname)
# load the csv in data frame
activity <- read.csv(fpath, sep = ",", as.is=TRUE)
head(activity)
library(dplyr)
activity<- tbl_df(activity)
```

remove the NAs of steps and group the date by date and calculate the totalsteps per day  
```{r}
activity<- activity[!is.na(activity$steps),]
activity_by_day<- group_by(activity, date)
sumsteps_by_day<- summarise(activity_by_day, totalsteps=sum(steps))
```

Histogram of the total number of steps taken each day and calculate Mean and median number of steps taken each day
```{r}
hist(sumsteps_by_day$totalsteps, xlab = "Total steps by day", main = "Histogram of total steps by day")
meansteps<- mean(sumsteps_by_day$totalsteps)
mediansteps<- median(sumsteps_by_day$totalsteps)
```
The mean steps per day is `r meansteps`, the median steps per day is `r mediansteps`

make the time series plot of the average number of steps taken
```{r}
activity_by_interval<- group_by(activity, interval)
meansteps_by_interval<- summarise(activity_by_interval, meansteps=mean(steps))
plot(meansteps_by_interval$interval, meansteps_by_interval$meansteps, type = "l", 
     lty=1, lwd=2, col="blue", xlab = "Interval(minutes)", ylab = "Steps", main = "Time series plot of the average number of steps taken")
```

Imputing missing data by mean value
```{r}
activityall<- read.csv(fpath, sep = ",", as.is=TRUE)
## Calculate the total number of missing values in the dataset (i.e. the total number of rows with ð™½ð™°s)
activityinputed<- activityall
for(i in 1:nrow(activityinputed)){
        if(is.na(activityinputed$steps[i])){
             interval_index<- activityinputed$interval[i]   
             row_index<- which(meansteps_by_interval$interval==interval_index)
             activityinputed$steps[i]<- meansteps_by_interval$meansteps[row_index]
        }
}
```

Group the data by date and calculate the totalsteps per day after NAs of steps were inputed
```{r}
activityinputed_by_day<- group_by(activityinputed, date)
sumstepsinputed_by_day<- summarise(activityinputed_by_day, totalsteps=sum(steps))
## plot the total steps afte missing values were inputed
hist(sumstepsinputed_by_day$totalsteps, xlab = "Total steps by day", main = "Histogram of total steps by day after missing values inputed")
```

transfer to date format and discriminate the data during weekday and weekend
```{r}
activityinputed$date<- as.Date(activityinputed$date, "%Y-%m-%d")
activityinputed$weekday<- weekdays(activityinputed$date, abbreviate = FALSE)
activityinputed$type_of_day[(activityinputed$weekday %in% c("æ˜ŸæœŸå…­", "æ˜ŸæœŸæ—¥"))]<- "Weekend" 
activityinputed$type_of_day[!(activityinputed$weekday %in% c("æ˜ŸæœŸå…­", "æ˜ŸæœŸæ—¥"))]<- "Weekday" 
```

make the Panel plot comparing the average number of steps between weekdays and weekends
```{r}
averagesteps<- aggregate(steps~type_of_day+interval, data = activityinputed, mean)
library(ggplot2)
qplot(interval, 
      steps, 
      data=averagesteps, 
      geom = "line",
      facets=type_of_day~., 
      xlab="Interval(minutes)", 
      ylab="Average Steps", 
      main="Comparison of average steps between weekdays and weekend",
      col=type_of_day)
```


